---
title: "a3_task4_logan_babcock.Rmd"
author: "Logan Babcock"
date: "10/27/2021"
output: html_document
---

### Introduction
This is the code for the analysis of urchin size data taken from various sites from 2001-2012, but here we are specifically looking at data for the 2010 Naples Reef treatment site for the Purple Urchin Species. 

*Data Citation:* Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages

library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
# read in data

urchins <- read.csv(here("data", "urchins.csv"))

```
### Wrangling the data
Code for organizing the data and making analysis easier.
```{r}
# wranging the data

# cleaning urchins and changing date class

purple_urchins_clean <- clean_names(urchins)

purple_urchins_clean %>% 
  mutate(date_new = ymd(date)) %>%  # change the date class and remove old column
  mutate(year = year(date_new)) %>% # new colum with only the year
  filter(common_name == "Purple Urchin", year == 2010, site == "NAPL") # keep purple urchins for NAPL site in 2010

```
### Exploring the data
here we use exploratory (informal and informational) graphs to explore the distribution of the data, and make observations about trends that are revealed. 
```{r}
 # making a exploratory graphs

# exploratory histogram

ggplot(purple_urchins_clean, aes(x = size)) + 
  geom_histogram(bins = 15) +
  facet_wrap("treatment") # separating the graphs by treatment

# exploratory qq plot

ggplot(purple_urchins_clean, aes(sample = size)) + # making the qq plot based on urchin size
  geom_qq() +
  facet_wrap(~treatment) # separating by treatment 
 
ggplot(purple_urchins_clean, aes(sample = size)) + # making the qq plot based on urchin size, with both populations considered together
  geom_qq()
```
This histogram shows us that the distributions are relatively similar, with the majority of sizes falling between 2 and 7 centimeters. The qqplot shows a straight line when both sets of data came from the normal distributions, comparing each set of data to see if they are from the same distribution. Here, the line is lower at the beginning of the graph, and slightly tapers off towards the end on the right side, this means that the distributions likely peaked in the middle compared to a normal-curve distribution. However, this trend is seen in both treatments of urchin population, meaning that each population likely follows this distribution.

### Summary Table
Summary statistics on the Purple Urchins using code to generate a table.

```{r}
# summary stats table 
purple_urchins_stats <- purple_urchins_clean %>%
  group_by(treatment) %>% # separating by treatment
  summarize( # summarizing main data points
    mean_size = mean(size),
    sd_size = sd(size),
    sample_size = n())

# finalizing stats table

kable(purple_urchins_stats, caption = "Table 1: Statistics on Purple Urchin Size by treatment at Naples Reef site", digits = 3, align = "c")
```

