---
title: "a3_task3_logan_babcock"
author: "Logan Babcock"
date: "October 26, 2021"
output: html_document
---

### Introduction 
This R makrdown document tracks the analysis of urchins data. 

*Data Citation:* Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
library(effsize)

# read in the data
urchins <- read_csv(here("data", "urchins.csv"))

# cleaning the data

urchins_clean <- clean_names(urchins)

```

### Section 1: Size comparisons of purple and red urchins
This section compares the sizes of the purple and red urchins, using data from all sites and all treatments. Figure 1 is a histogram visualizing the differences in urchin size distribution based on species.
```{r}
# creating a faceted histogram
ggplot(data = urchins_clean, aes(x = size)) + # histograms only need one axis
  geom_histogram(bins = 15) + # making the bin size smaller to better show the data
  facet_wrap(~ common_name) + # separating the data by urchin type
  labs(x = "Urchin diameter (cm)", #adding x and y axis labels
       y = "Urchin count", # adding figure title
       title = "Urchin size distribution by species (Babcock)",
       caption = "Figure 1: Histogram showing the distribution of Urchin size measured by diameter in cm. Shown for both purple and red urchin species.") + # adding figure caption
  theme_minimal() 
  
```

```{r}
# creating a summary statistics table

urchins_summary <- urchins_clean %>% # start with urchins data
  group_by(common_name) %>% # group by species 
  summarize(sd(size), mean(size), median(size), n()) %>%  # summary table values wanted

```

```{r}

```

